Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Using Expert Knowledge in Model Construction

### 2024-08-06 paper/main.tex
Updates abstract

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	Causal Discovery has received substantial attention in the DAG
	literature and many automated algorithms have been proposed for
	constructing DAGs from data. However, the application of these
	algorithms in applied fields remain limited due to challenges in using
	them in practice for example they make obvious mistakes, output CPDAGs
	that need to be oriented using expert knowledge before estimation
	methods can be applied. As a result of these challenges, researchers
	typically construct these model by hand. In this paper, inspired by
	specification search in SEMs where the method suggests changes in the
	model that improves the fit the most, we present a method based on
	Conditional Independence testing and a novel measure of partial
	association. The method iteratively suggests addition and removal of
	edges between pairs of variables, while the user can choose the
	direction of the edge. Using empirical analysis we show that this
	manual construction approach can outperform automated algorithms if the
	user gets the direction correct in <x out of y> cases. We also show
	that the user expert can be replaced with an LLM.
\end{abstract}

>>>>

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Using Expert Knowledge in Model Construction}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	Causal Discovery has received significant attention in the Directed
	Acyclic Graphs (DAGs) literature, leading to the development of
	numerous automated algorithms for constructing DAGs from data. Despite
	these advancements, the application of these algorithms in applied
	domains remain limited, with researchers often preferring to construct
	DAGs manually based on their expert knowledge. We hypothesize that the
	reluctance to use automated algorithms arise due to issues with these
	algorithms such as their tendency to make obvious mistakes, and
	outputting the Markov Equivalence Class (MEC) instead of a DAG, which
	is often required for further analysis. Even when using these
	algorithms, researchers have to rely on their expert knowledge to make
	modifications to the learned model. Given this reliance on manual model
	construction and modification, in this paper, we introduce a method to
	assist researchers in this task inspired by specification search in
	Structural Equation Models(SEMs). We present an iterative approach to
	model modification utilizing Conditional Independence (CI) testing
	combined with a novel measure of partial association for mixed data to
	rank potential additions or removal of edges to the DAG such that the
	improvement in the model's explanation of the data is improved the
	most. Specifically, the method ranks violations of implied conditional
	independences (CI) in the data using the measure of partial
	association. The user can choose to fix these violations by adding
	edges between the pair of variables while using their domain expertise
	to decide the direction of edge between the variables. Through
	empirical analysis, we demonstrate that taking a greedy approach (i.e.
	fixing the worst ranked violation first) with this manual construction
	approach can get comparable performance to automated algorithms if the
	user are able to get the edge direction correctly between the variables
	at least $ 1 $ out of $ 3 $ prompts (check if this is correct).
	Alternatively, we show that a Large Language Model (LLM) can also be
	used to help in deciding the direction between the variables.
\end{abstract}
